---
title: 'Statistical Inference : ToothGrowth'
author: "courserans"
date: "February 20, 2015"
output:
  html_document:
    keep_md: yes
---
#### Overview
In this project we will investigate the ToothGrowth Data set with techniques learnt in the Coursera Statistical Inference class. The ToothGrowth data set documents the effect of Vitamin C on tooth growth in Guinea Pigs. The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

#### Exploratory Data Analysis
* Plot
```{r, fig.align="center",fig.height=4.5, fig.width=5.5, fig.path='fig/'}
library(ggplot2)
library(grid)
library(gridExtra)
g1 <- ggplot(ToothGrowth, aes(factor(dose), len)) + geom_boxplot((aes(fill = factor(dose)))) + facet_grid(. ~ supp) + theme_bw()
g2 <- ggplot(ToothGrowth, aes(factor(supp), len)) + geom_boxplot((aes(fill = factor(supp)))) + facet_grid(. ~ dose) + theme_bw()
grid.arrange(g1,g2,nrow=2)
```
Fig1 : As the top panel of plot indicates, there seems to be evidence of dose levels influencing the length of the tooth growth for a given delivery method. However the influence of delivery method for a given dose level is not so clear.

* Contingency Table
```{r, comment=""}
table(ToothGrowth$supp,ToothGrowth$dose)
```
The total no of observations is `r nrow(ToothGrowth)`

####  Basic Summary of Data
```{r, comment="" }
library(plyr)
print(ddply(ToothGrowth,.(supp),summarize,mean=mean(len),sd=sd(len),n=length(len)),row.names=FALSE,justify="centre")
print(ddply(ToothGrowth,.(dose),summarize,mean=mean(len),sd=sd(len),n=length(len)),row.names=FALSE,justify="centre")
```
#### Confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose.
- Hypothesis testing for difference of means by supp("Delivery Method : OJ,VC")
```{r}
tsupp <- t.test(len~supp,ToothGrowth)
tsupp
```
###### Result : Since p = 0.06, we cannot reject the null hypothesis. 
- Hypothesis testing for difference of means by dose ("Dose Levels : 0.5,1,2")

```{r}
tdose0.51.0 <- t.test(len~dose,ToothGrowth,subset=((dose==0.5)|(dose==1)))
tdose1.02.0 <- t.test(len~dose,ToothGrowth,subset=((dose==1.0)|(dose==2.0)))
tdose2.00.5 <- t.test(len~dose,ToothGrowth,subset=((dose==2.0)|(dose==0.5)))
tdose0.51.0
tdose1.02.0$p.value  ## P value of the t.test for Dose 1.0 and 2.0 mean comparison. For brevity just printed the p values.
tdose2.00.5$p.value  ## P value of the t.test for Dose 2.0 and 0.5 mean comparison. For brevity just printed the p values.
```
###### Result : Based on the p values, we can reject the null hypothesis that the means are the same across the dose levels(0.5,1,2).

##### Conclusion & Assumption
###### Based on the hypothesis testing we can conclude that dose levels has effect on tooth growth. We also observe that increasing the dose levels increases the tooth growth. However supp(Delivery Method) has no effect on the tooth growth. The hypothesis testing results are valid under the assumption that a) observations are independent b) variance of the groups is not equal (var.equal=FALSE) c) The confidence level of the interval = 95%.d) Not a paired test (paired = FALSE)
